\documentclass[a4paper,12pt,english,final]{epsc_tfc_pfc}
%% a4paper: mida paper. No tocar
%% 12pt: mida de la font. No tocar

%%  - OPCIONS A CONFIGURAR:
%%     - Estat del document: final o draft
%%       NOTA: Draft no inserta les figures i indica quan el text sobrepassa els marges.

%%  - IDIOMES QUE S'USARAN EN EL DOCUMENT: catalan, spanish, english, french...
%%    NOTA: per canviar d'idioma al mig del document usar:
%%          \selectlanguage{nom_idioma}
\usepackage[english,catalan]{babel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%% PAQUETS LATEX RECOMANABLES A UTILITZAR
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% NOTA: es possible que algunes distribuicions Linux o Windows 
%%%       no portin aquests paquets instal·lats per defecte.

%% El paquet isolatin1 és extramadament útil. 
%% Permet escriure els accents directament amb l'editor de texte
%% sense haver de fer coses com per exemple: introducci\'o
\usepackage[latin1]{inputenc}             

%% Símbols matemàtics de la American Mathematical Society
\usepackage{amssymb,amsmath,amsfonts}

\usepackage{titlesec}

%% Glossaries
%\usepackage[nonumberlist,section=chapter]{glossaries}
\usepackage[nonumberlist]{glossaries}
\makeglossaries

%% Figures wrapped on text
\usepackage{wrapfig}

%% El paquet array proporciona eines molt útils a l'hora de fer 
%% equacions amb matrius
\usepackage{array}             

%% Permet fer taules fusionant cel·les de files consecutives
%\usepackage{multirow}          

%% Permet canviar els colors del document
%\usepackage{color,colortbl}
\usepackage{float}

\restylefloat{table}

\usepackage{textcomp}

\usepackage[export]{adjustbox}

\usepackage{times}

%\url
\usepackage{url}

\begin{document}
\pagestyle{empty}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% IDIOMA PRINCIPAL DEL DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\selectlanguage{english}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% PORTADA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Portada generada automàticament a partir del fitxer de dades
\portada

%% RESUM
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% NOTA: la longitud passada com a parametre d'entrada 
%       s'ha d'ajustar ``a ull'' fins que el requadre del resum ocupi tota la pagina 
\begin{resum}{1cm}
Cada vegada més l'Internet de les Coses està en auge impulsant tot tipus de dispositius incrustats i sistemes operatius per aquests dispositius. L'objectiu d'aquest projecte de fi de carrera és desenvolupar una llibreria capaç d'executar aplicacions desenvolupades per a .NET Micro Framework (màquina virtual per a dispositius incrustats) en dispositius que puguin executar qualsevol versió de Linux. La idea és que utilitzant la llibreria desenvolupada es pugui executar qualsevol aplicació de .NET Micro Framework a un ordinador normal.
\\
\\
En aquest document es presenta el plantejament i el desenvolupament d'IOSharp que correspon a la implementació en C\# del sistema comentat anteriorment. Un cop aquesta implementació s'ha dut a terme, s'ha passat a la fase de proves amb les aplicacions existents on s'ha pogut determinar que el rendiment de la llibreria no és òptim. En aquest punt és on hi ha hagut una de les altres contribucions principals en el projecte, s'ha participat en el desenvolupament d'una eina anomenada AlterNative capaç de traduir aplicacions escrites en .NET a C++, d'aquesta manera s'avaluarà la millora del rendiment de la llibreria al traduir-la amb l'AlterNative.
\end{resum}

%% OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% NOTA: la longitud passada com a parametre d'entrada 
%       s'ha d'ajustar ``a ull'' fins que el requadre del resum ocupi tota la pagina 
\begin{overview}{1cm}
Increasingly, the Internet of Things is promoting all types of embedded devices and operating systems for these devices. The aim of this project is to develop library capable of running applications developed for .NET Micro Framework (virtual machine for embedded devices) on devices that can run any version of Linux. The idea is that using this library a developer can execute any application developed for .NET Micro Framework in any computer.
\\
\\
This thesis presents the approach and the development of IOSharp which is an implementation in C\# of the system discussed above. Once this implementation has been done, the test phase has been started with existing applications. In this case it has been proved that the performance of this library is not optimal compared with the original one. At this point is where started another of the main contributions to the project, participating in the development of a tool called Alternative which is able to translate applications written in .NET to C++, so by this way it will be possible to analyse the improvement of this library after translating with AlterNative.
\end{overview}

%% DEDICATORIA (opcional)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{dedicatoria}
Escriure aquí opcionalment la dedicatòria.
\end{dedicatoria}

%% INDEX de continguts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty} 
\tableofcontents
\cleardoublepage

%% INDEX de figures (opcional, comentar les 3 linies si no es desitja)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}
\listoffigures
\cleardoublepage

%% INDEX de taules (opcional, comentar les 3 linies si no es desitja)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}
\listoftables
\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%                    INTRODUCCIÓ                          %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% NOTA: El text passat com a parametre d'entrada 
%%       és ''introducció'' amb l'idioma en que es redacti el projecte
\pagestyle{fancy} 

%% First include Glossary to be used in the following chapters
\include{chapters/glossary/glossary}

\begin{intro}{Introduction} 
Embedded systems have become more powerful over time passing from 8 bit controllers to 8 or 16 bit microprocessors or even 32 bit ARM microprocessors. Apart from the increase of the processing power, the memory included in this devices has also increased, from tens of Kilobytes to tens or hundreds of Megabytes. One of the reasons for these changes has been the price drop on production. These new embedded systems offer a performance similar to the computers from the nineties and most of them implement operating systems which helps on reducing the difficulty to create embedded applications. With an operating system running over the bare metal of the chip, the developer will get all the underlying hardware abstracted to different APIs and libraries and avoiding low level interaction with the hardware. They also offer interesting features such as memory control and allocation, threading or dynamic program loading. In example, one of the top notch devices nowadays is the Raspberry Pi which mounts an ARMv6, 512MB of RAM and some I/O features such as GPIOs, SPI, UART. This hardware can be equivalent to a nineties computer similar to a Pentium II so they are really powerful to the task that they may do.
\\
One of the operating systems for embedded devices is .NET Micro Framework (NETMF) developed by Microsoft. This system is the smallest version of .NET Framework and is oriented to resource-constrained devices for embedded applications. This system offers different communication protocols and methods like the \gls{GPIO} ports, the \gls{SPI}, \gls{UART} and \gls{I2C}. There is no official implementation or port of Micro Framework capable of running on standard computers (a normal desktop), so any application written for this operating system will not work on Linux or Windows. Although a minimal port of Micro Framework exists for a Linux board called Eddy, it does not offer all the hardware features that Micro Framework does.
\\
The aim of IOSharp is to solve this lack and get Micro Framework applications run on any Linux machine which is capable of running applications using the .NET Framework (the complete stack designed by Microsoft). So this project instead of writing a complete port of the Micro Framework runtime to run on Linux, is an extension to the classes provided by the .NET Framework. Basically, IOSharp offers the IO functions, methods and classes that are missing on .NET Framework, and although the implementation of the classes is different in IOSharp than in Micro Framework, the namespaces, methods, class naming, etc is equal to the original one so this makes much easier to migrate between Micro Framework to .NET Framework with IOSharp. 
\\
The reason for this implementation is trying to migrate the Micro Framework code that runs on a Netduino Mini which is the gateway of a Wireless Sensor Network (\gls{WSN}). One of the problems that this platform has is that the actual device is getting out of system resources, so in order to keep the existing code, different solutions are being researched. In this case IOSharp intends to be one of those solutions achieving the deployment of this WSN gateway software on a Raspberry Pi which runs a Linux operating system.
\\
\\
After achieving the first goal, there is a second evolution of this project which consists on extending a code translating tool called AlterNative which is being developed by Alex Albalá and Juan López. This translator is capable to get the source code from a C\# assembly and then translate it to C++ trying to improve the performance. C++ theoretically performs better than C\# applications, but normally are more platform oriented, so a Windows assembly will not work on a Linux system, but one of the interesting features of AlterNative is the generation of a highly portable code that can be compiled on (and for) any operating system, i.e. Windows, Linux, MacOSX, iOS, Android, etc. Another interesting point of this tool is that the generated code is similar to C\# so a developer used to this language will be able to understand or even write programs using the translated code.
\end{intro}

\pagestyle{fancy} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% INCLOURE A PARTIR D'AQUÍ TOTS ELS CAPÍTOLS DE LA MEMORIA   %%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% First include Glossary to be used in the following chapters
%\include{chapters/glossary/glossary}

%% Include the chapters
\include{chapters/proposal/proposal}
\include{chapters/state_of_the_art/state_of_the_art}
\include{chapters/iosharp/iosharp}
\include{chapters/iosharp-examples/iosharp-examples}
\include{chapters/alternative/alternative}
\include{chapters/performance-tests/performance-tests}
\include{chapters/conclusions/conclusions}
%% Remove the final dot and print the glossaries
%\renewcommand*{\glspostdescription}{}
\printglossaries

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%  BIBLIOGRAFIA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Per la bibliografia hi ha 2 opcions: generarla amb la utilitat BibTeX 
%%%                                      o fer-la ''a ma''
%%% NOTA: podeu trobar facilment informació sobre BibTeX a:
%%%  http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/


%%% OPCIO 1: BibTeX -> descomentar les dues línies
%%% a)  Estil de bibliografia\\
%\bibliographystyle{unsrt}  
%%% b) Indicar els fitxers que contenen la bibliografia
%\bibliography{bibliography2}  

%%% OPCIO 2: bibliografia manual
%%%
%%% L'argument d'entrada es el numero de referencies que s'inclouen
\renewcommand{\bibname}{REFERENCES}
\begin{thebibliography}{9}

%% Llibres:  Autor/s (cognoms i inicials dels noms), títol del llibre (en cursiva), editor, ciutat i any de publicació. Quan es cita el capítol d'un llibre s'ha d'indicar el títol del capítol (entre cometes), el títol del llibre (en cursiva) i els números de pàgines amb la primera i la darrera incloses.
%%  Exemple de capitol en llibre
\bibitem{cite:embedded-linux}
Weinberg, B., {\it Uniting Mobile Linux Application Platforms}, LinuxPundit.com, 2008. Also available at \url{http://www.linuxpundit.com/cv/docs/Platforms_WP_LP.pdf}

\bibitem{cite:sensors-everywhere}
Gómez, C., Paradells, J. and E. Caballero, J, "Operating Systems", {\it Sensors Everywhere. Wireless Network Technologies and Solutions}, p.273-278, Fundación Vodafone España, 2010. Also available at \url{http://fundacion.vodafone.es/static/fichero/pre_ucm_mgmt_002618.pdf}

\bibitem{cite:NETMF}
Kühner J., {\it EXPERT .NET MICRO FRAMEWORK}, Apress, United States of America, 2010.

\bibitem{cite:delegates}
Kiely, D., "Delegates Tutorial", {\it The Microsoft Developer Network (MSDN)}. Available at \url{http://msdn.microsoft.com/en-us/library/aa288459(v=vs.71).aspx}

\bibitem{cite:ILSpy}ILSpy homepage, \url{http://ilspy.net/}, 2013

\bibitem{cite:compilers1}
A. V. Aho, R. Sethi, J. D. Ullman, {\it Compilers. Principles, Techniques, and Tools}, Bell Telephone Laboratories, United States of America, 1986. 

\bibitem{cite:compilers2}
D. Grune, H. E. Bal, C. J.H. Jacobs, K. G. Langendoen, {\it Modern Compiler Design}, Vrije Universiteit, Amsterdam and Delf University, John Wiley \& Sons, Ltd, England, 2000

%\bibitem{cite:oom}
%Khan, S., Qureshi, K. and Rashid, H., "Performance Comparison of ICE, HORB, CORBA and Dot 
%NET Remoting Middleware Technologies", {\it International Journal of Computer Applications}, 3(11), 15-18 (2010). Also available at \url{http://www.ijcaonline.org/volume3/number11/pxc3871105.pdf}

%\bibitem{cite:marea}
%López, J., Royo, P., Barrado, C., Pastor, E., "Applying marea middleware to UAS communications", {\it  In Proceedings of the AIAA Infotech@Aerospace Conference and AIAA Unmanned Unlimited Conference 2009}, Seattle (WH). Also available at \url{http://upcommons.upc.edu/e-prints/bitstream/2117/9248/1/infotech09.pdf}

%\bibitem{cite:thesis-soa-avionics}
%López, J., "Service Oriented Architecture for Embedded (Avionics) Applications", {\it  The PhD Program on Computer Architecture Technical School of Castelldefels Technical University of Catalonia}, Barcelona, 2011. Also available at \url{https://dl.dropbox.com/u/2857619/thesis-small.pdf}

%\bibitem{cite:delegate}  
%Kiely, D., "Delegates Tutorial" in {\it The Microsoft Developer Network (MSDN)}. Available at \url{http://msdn.microsoft.com/en-us/library/aa288459(v=vs.71).aspx}

%\bibitem{cite:serialization} 
%Albahari, J. and Albahari B., "Serialization", Chap. 17 in {\it C\# 5.0 in a Nutshell: The definitive reference},
%O`REILLY, Roumeliotis, R., pp. 691-728, Sebastopol (CA), 2012.

%\bibitem{cite:performance-sockets} 
%Kiely, D., "Get Closer to the Wire with High-Performance Sockets in .NET" in {\it The Microsoft Developer Network (MSDN) Magazine}. Available at \url{http://msdn.microsoft.com/es-es/magazine/cc300760(en-us).aspx}

%\bibitem{cite:backporting}
%Books Llc, Source Wikipedia, "Software Quality: Software Crisis, Kludge, Second-System Effect, Workaround, Reliability Engineering, Fault-Tolerant System", Books Llc, Memphis (Tennessee), 2011.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%                           APENDIXS                         %%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{empty}  % no tocar

%% Descomentar una de les dues línies següents, en funció de:
%%  a) els apendixs s'encuadernaran apart (amb portada) 
%%  b) els apendixs s'enquadernen amb el mateix projecte (sense portada). 
%% Recordeu que si tot el document (amb apèndixs) excedeix les 100 pagines 
%% s'ha d'enquadernar a part
\appendix\ambportada
%\appendix\senseportada


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% INCLOURE A PARTIR D'AQUI TOTS ELS CAPÍTOLS DELS APENDIXS   %%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{chapters/appendices/tools}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% i  aixo es tot! ;)
\cleardoublepage
\end{document}